Comparison of several stochastic and deterministic derivative-free global optimization algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}

### 2019-02-11 2019/MOTOR_Sovrasov/motor_paper.tex
MOTOR: update the paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{subfig}
\usepackage{cite}

@@

\begin{document}
%
\title{Comparison of several derivative-free global optimization algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Vladislav Sovrasov}
%
\authorrunning{V. Sovrasov}
%
\institute{Lobachevsky State University of Nizhni Novgorod, Russia \\
  \email{sovrasov.vlad@gmail.com}
}%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
This work considers several stochastic and deterministic deri-vative-free global optimization
algorithms. In the first part of the paper popular sequential open-source solvers are compared against
Globalizer solver, which is developed at the Lobachevsky State University. The Globalizer is
designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
competitive performance with other similar solvers. The comparison is done on several sets of
challenging multi-extremal benchmark functions. The second part of this work is devoted to
comparison between the Globalizer and MIDACO solvers on systems with shared and distributed
memory. MIDACO is a state-of-the-art global solver included to the TOMLAB optimization
environment for MATLAB. Results of the benchmark show advantages of the Globalizer on small-dimensional, but sufficiently multi-extremal benchmark functions.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{deterministic global optimization $\cdot$ stochastic global optimization
  $\cdot$ parallel numerical methods $\cdot$ derivative-free algorithms $\cdot$ black-box
optimization}
\end{abstract}
%
\section{Introduction}

Nonlinear global optimization of non-convex functions is considered to be one of the most
difficult problems of mathematical programming. The finding of the global minimum of
a function of several variables often appears to be more difficult than local optimization in a
thousand-dimensional space. In local optimization applying of the simplest gradient descent
method may appear to be enough, whereas in order to \textit{guarantee} finding the global
optimum, an optimization method has to accumulate information about behavior of the
objective function within the whole search domain
\cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. Nowadays, various stochastic global
optimization algorithms have attracted much attention, first of all, the evolution ones
\cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}. They have rather simple structure
and allow solving the problems of large dimensionality. However, the evolution algorithms ensure
the global convergence in the probabilistic sense only.

In the present paper, the open-source implementations of nine different global optimization methods
were considered. All the algorithms were tested using a set of 900 essentially multiextremal functions,
which has been generated using generators of test problems \cite{Gaviano2003, grishaginClass}.
Besides the comparison of the sequential algorithms, a comparison of the MIDACO solver
\cite{Schlueter2012} with the Globalizer software system \cite{globalizerSystem,Strongin2018} on a subset of 200 test problems has been
performed in the Lobachevsky supercomputer environment.

@@

Taking into account the experience of the preceding works in the field of comparison of the
optimization methods, two generators of the test problems of different nature will be used in the
present work. Using these ones, 9 sets of 100 problems of various complexity with the
dimensionality varying from 2 to 5 were generated. Besides the comparison of the sequential
methods, a comparison of the efficiencies of two parallel algorithms is presented in the work as well.

@@

\subsection{Parallel Algorithm of Global Search}
\subsubsection{Dimension Reduction with Space-Filling Curves}
Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or \textit{evolvents}) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction \cite{sergeyevStronginLera2013, strongin1978, strSergGO}.
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the % minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
\end{equation}
where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem
(\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
given in \cite{strSergGO}.

@@

\subsubsection{Algorithm of Global Search on Shared Memory}
\label{sub:ags}
Parallel optimization methods applied in Globalizer to solve the reduced problem
(\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows ---
see \cite{strongin1978}, \cite{strSergGO}.
\par
The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the
following rules.

\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of
increasing value of coordinates
$0=x_0<x_1<...<x_{k+1}=1$.

\textit{Rule 2}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq
k+1\).

\textit{Rule 3}. Determine the \(p\) intervals with the maximum characteristics $R(t_j)=\max_{1\leq i
\leq k+1}R(i),\: j=\overline{1,p}$.

\textit{Rule 4}. Execute new trials at points \(x^{k+j},\: j=\overline{1,p}\) located within intervals
with
maximum characteristics from the previous step
  $x^{k+j}=d(x_{t_j}),\: j=\overline{1,p}$.

The stopping condition, which terminated the trials, is defined by the inequality
$\rho_{t_j}<\varepsilon,\: j=\overline{1,p}$
for the intervals with maximum characteristics from Step 3 and \(\varepsilon >0\) is the
predefined accuracy of the optimization problem solution. If the stopping condition is not satisfied,
the index \(k\) is incremented by \(p\), and the new global optimization iteration is executed.

This method is employed in Globalizer to organize parallel computations on shared memory: each of
\(p\)
trials can be carried out on one of \(p\) local computation units.

The convergence conditions and exact formulas for decision rules $R(i)$ and $d(x)$ of the
described algorithm are given, for example, in \cite{strSergGO}.

\subsection{Sequential Methods}
\begin{itemize}
  \item \textbf{Algorithm of Global Search}. Sequential version of the method described in Section
\ref{sub:ags}.

  \item \textbf{Locally-biased Algorithm of Global Search (AGS\(l\))} \cite{indexMethod}. It's a
modification of
  the original AGS which make it more locally oriented by alternately using of two types of
characteristics in the Rule 2 from the Section \ref{sub:ags}.

  \item \textbf{Multi Level Single Linkage} \cite{Kan1987StochasticGO}. MLSL is an improved
multistart algorithm.
  It samples low-discrepancy starting points and does local optimizations from them. In contrast to
the dummy multistart schemes
  MLSL uses some clustering heuristics to avoid multiple local descents to already explored local
minima.

@@

All the mentioned algorithms are available in source codes as parts of wide-spread optimization packages (except of AGS and AGS\(l\), which are implemented in the Globalizer system).
DIRECT, DIRECT$l$, CRS, MLSL and StoGO are part of the NLOpt library \cite{nlopt}.
Differential Evolution and DSA can be found in
the latest version of the SciPy \cite{scipy} package for Python.

@@

The results of various algorithms on different problem classes depend on the adjustments of
algorithms directly. In most cases, the authors of software implementations are oriented onto the
problems of medium difficulty. In order to obtain a satisfactory result when solving the essentially
multiextremal problems, a correction of some parameters is required. When conducting the
comparison, the following parameters for the methods were employed:
\begin{itemize}
  \item in the AGS\(l\) method, the parameter of alternation the
global and local iterations was set to be equal to 5:1;
  \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
  \item in the SDA method, the parameter \(visit=2.72\).
\end{itemize}

The rest parameters were varied subject to the problem class (see Table \ref{tab:params}).

@@

The results of running the optimization methods on the considered problem classes are presented in
Tables \ref{tab:trials}, \ref{tab:solved}. The DIRECT and AGS\(l\) methods have demonstrated the
best convergence speed on all classes, at that AGS\(l\) inferior to DIRECT on the problems from the
Simple classes and has an advantage on the problems of the Hard classes. As one can see from Table
\ref{tab:solved}, the deterministic methods (AGS, AGS\(l\), DIRECT, and DIRECT\(l\)) were the
most reliable. Among the stochastic methods, MLSL and SDA have demonstrated the highest
reliability.

\begin{table}
\begin{center}
\caption{Averaged number of trials executed by sequential methods for solving the test
optimization problems}
\resizebox{\textwidth}{!}{%
  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
    \hline
    & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
  \hline
  \(F_{GR}\)     & 193.1 &  \textbf{158.3} & 400.3 & 182.3 & 214.9 & 947.2 & 691.2 & 1257.3 &
1336.8 \\
  \hline
  GKLS 2d Simple &  254.9 & 217.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2
& 1251.5 \\
  \hline
  GKLS 2d Hard   &  728.7 & \textbf{488.0} & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 &
1041.1 & 2532.2 \\
  \hline
  GKLS 3d Simple &  1372.1 & 1195.3 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 &
5956.94 & 3856.1 \\
  \hline
  GKLS 3d Hard   &  3636.1 & \textbf{1930.5} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 &
6914.3 & 7843.2 \\
  \hline
  GKLS 4d Simple &  26654.1 & 11095.7 & 37436.8 & \textbf{7824.3} & 15994.1 & 41514.3 &
21417.9 & 19157.7 & 59895.4 \\
  \hline
  GKLS 4d Hard   &  54536.8 &  \textbf{23167.8} &  73779.3 &  23204.4 &  54489.9 &  80247.2
&  68815.5 &  27466.1 &  109328.1  \\
  \hline
  GKLS 5d Simple &  29810.0 & 11529.0 & 143575.0 & \textbf{7166.5} & 13970.5 & 52647.6 &
34255.3 & 73074.5 & 91580.4 \\
  \hline
  GKLS 5d Hard   &  113129.1 & 67652.7 & 165192.8 & \textbf{66327.4} & 164390.6 & 138766.2
& 116973.1 & 105496.9 & 155123.8 \\
  \hline
\end{tabular}}
  \label{tab:trials}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Number of test optimization problems solved by sequential methods}
  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
    \hline
    & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
  \hline
  \(F_{GR}\)     &  100 & 100 & 76  & 100 & 100 & 97  & 96  & 96  & 67\\
  \hline
  GKLS 2d Simple &  100 & 100 & 85  & 100 & 100 & 100 & 100 & 98  & 90\\
  \hline
  GKLS 2d Hard   &  100 & 100 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
  \hline
  GKLS 3d Simple &  100 & 97  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
  \hline
  GKLS 3d Hard   &  100  & 99   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
  \hline
  GKLS 4d Simple &  100 & 100 & 46  & 100 & 100 & 94  & 78  & 59  & 16 \\
  \hline
  GKLS 4d Hard   &  100 & 100 & 47  & 99  & 97  & 94  & 72  & 32  & 10  \\
  \hline
  GKLS 5d Simple &  100 & 100 & 68  & 100 & 100 & 98  & 100 & 77  & 9  \\
  \hline
  GKLS 5d Hard   &  97  & 99  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
  \hline
  \end{tabular}
  \label{tab:solved}
\end{center}
\end{table}

@@

In the present paper, several sequential and parallel global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
making the following conclusions:
\begin{itemize}
  \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems;
  \item the parallel version of the Globalizer system demonstrates good speedup
when running on several nodes, on each of which a single objective function value per iterations is
computed. When solving the problems with fast computable objective functions and using multiple
threads on the nodes, the speedup for the Globalizer system degrade with increasing
number of nodes;
  \item the MIDACO system is the most suitable for simple global optimization problems with the
fast-computable objective functions. In this case, MIDACO is reliable enough and provides a linear
speedup with increasing number of nodes and threads executed on these ones in parallel.
\end{itemize}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{subfig}
\usepackage{cite}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\sign}{sign}

%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}

@@

\begin{document}
%
\title{Comparison of several stochastic and deterministic derivative-free global optimization algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Vladislav Sovrasov}
%
\authorrunning{V. Sovrasov}
%
\institute{Lobachevsky State University of Nizhni Novgorod, Russia \\
  \email{sovrasov.vlad@gmail.com}
}%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
In this paper popular sequential open-source solvers are compared against
Globalizer solver, which is developed at the Lobachevsky State University.
The Globalizer is designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
competitive performance with other similar solvers. The comparison is done on several sets of
challenging multi-extremal benchmark functions. Also this work considers a method of
heuristic hyperparameters control for Globalizer allowing to reduce amount of
initial tuning before optimization. The proposed scheme allows rapidly increase convergence speed
of Globalizer by switching between ``local'' and ``global'' search phases in runtime.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{deterministic global optimization $\cdot$ stochastic global optimization
  $\cdot$ parallel numerical methods $\cdot$ derivative-free algorithms $\cdot$ black-box
optimization}
\end{abstract}
%
\section{Introduction}

Задача отысканя глобального минимума нелинейных невыпуклых функций традиционно считается одной из самых трудных
задач математического программирования. Зачастую она оказывается сложнее, чем локальная оптимизация в существенно многомерном пространстве.
Для последней может оказаться достаточно применения простейшего метода градиентного спуска или pattern search algorithms \cite{torczon1997},
в то время как чтобы \textit{гаранитрованно} отыскать глобальный оптимум методам
оптимизации приходится накапливать информацию о поведении целевой функции во всей области поиска \cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. В последнее время стали популярны
различные стохастические алгоритмы глобально оптимизации, прежде всего эволюционные \cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}. Они имеют довольно простую структуру, позволяют решать задачи большой размерности, однако обеспечивают глобальную сходимость только в вероятностном смысле.

В данной работе рассмотрены open-source реализации восьми различных методов глобальной оптимизации, представленные
в библиотеке NLOpt\cite{nlopt} и пакете SciPY\cite{scipy}.
Все алгоритмы были протестированы на наборе из 900 существенно многоэкстремальных функций, который был сгенерирован с
помощью специализированных генераторов задач \cite{Gaviano2003, grishaginClass}.

@@

Taking into account the experience of the preceding works in the field of comparison of the
optimization methods, two generators of the test problems of different nature will be used in the
present work. Using these ones, 9 sets of 100 problems of various complexity with the
dimensionality varying from 2 to 5 were generated.

@@

\subsection{Algorithm of Global Search}
\label{sub:ags}
\subsubsection{Dimension Reduction with Space-Filling Curves}
Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or \textit{evolvents}) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction \cite{sergeyevStronginLera2013, strongin1978, strSergGO}.
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the % minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
\end{equation}
where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem
(\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
given in \cite{strSergGO}.

@@

Optimization method applied in Globalizer to solve the reduced problem
(\ref{eq:oneDimTask}) is based on the AGS method, which can be presented as follows ---
see \cite{strongin1978}, \cite{strSergGO}.
\par
The algorithm considered for solving the stated problem implies generating
a sequence of points \(x_k\), in which the values of the minimized function \(z_k = f(x_k)\)
are computed. Let us call the process of computating the function value
(including calculating an image \(y^k=y(x^k)\)) a trial, and the pair \((x^k,z^k)\) ---
the result of the trial. A set of the pairs \(\{(x^k,z^k)\}, 1\leqslant k\leqslant n\)
makes up the search information accumulated by the method after executing \(n\) steps.
\par
The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the
following rules.

\textit{Step 1.} Renumber the points in the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
which includes the boundary points of the interval \([0,1]\) as well as the points of
preceding trials, by the lower indices in order of increasing coordinate values  i.e.
\begin{displaymath}
0=x_0<x_1<\dotsc<x_{k+1}=1
\end{displaymath}
\par
\textit{Step 2.} Assuming \(z_i=f(x_i),1\leqslant i\leqslant k\), compute the values
\begin{equation}
\label{eq:step2}
\mu=\max_{1\leqslant i\leqslant k}\dfrac{|z_i-z_{i-1}|}{\Delta_i},
\begin{matrix}
    M =
    \left\{
    \begin{matrix}
    r\mu,\mu>0 \\
    1,\mu=0
    \end{matrix} \right.
    \end{matrix}
\end{equation}
where \(r>1\) is a predefined parameter for the method, and \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
\par
\textit{Step 3.} For each interval \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), compute the
characteristics according to the formulae
\begin{equation}
\label{step3_1}
R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
\end{equation}
\begin{equation}
\label{eq:step3_2}
R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1.
\end{equation}
\par
\textit{Step 4.} Determine the interval with the maximum characteristic \((x_{t-1}, x_t),\:t=\argmax_{1\leqslant i \leqslant k+1}R(i)\)

\textit{Step 5.} Execute a new trial at point \(x_{k+1}\) computed according to the formula
\begin{displaymath}
x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2},t=1,t=k+1,
\end{displaymath}
\begin{displaymath}
\label{step5}
x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,1<t<k+1.
\end{displaymath}

The stopping condition, which terminated the trials, is defined by the inequality
\(\Delta_{t}\leqslant \varepsilon\)
for the interval with the maximum characteristics from Step 4 and \(\varepsilon >0\) is the
predefined accuracy of the optimization problem solution. If the stopping condition is not satisfied,
the index \(k\) is incremented by \(1\), and the new global optimization iteration is executed.

The convergence conditions of the described algorithm are given, for example, in \cite{strSergGO}.

\subsubsection{Hyperparameters tuning}
Параметр $r$ из (\ref{eq:step2}) непосредственно влияет на глобальную сходимость AGS (see \cite{strSergGO}, Chapter 8):
при достаточно большом значении $r$ метод гарантированно сходится ко всем глобальным минимумам целевой функции.
В то же время, согласно (\ref{eq:step3_2}), при бесконечно большом значении $r$ AGS превращается в полный перебор
по равномерной сетке.

В идеальном случае для обеспечения наибольшей скорости сходимости оценка константы Липшица из (\ref{eq:step2})
не должна быть слишком завышенной, но на практике реальное значение $L$ из (\ref{eq:lip}) неизвестно и приходится
либо брать заведомо большое значение $r$, либо проводить несколько запусков AGS с разными параметрами. Чтобы в некотрой
степени решить обозначенную проблему выбора $r$ будем использовать следующую схему:
\begin{itemize}
  \item совершить $q$ итераций AGS при $r=r_{max}$;
  \item совершить $q$ итераций AGS при $r=r_{min}$;
  \item повторять предыдущие шаги до схоимости или исчерпания лимита итараций.
\end{itemize}

В указанном алгоритме $r_{min} < r_{max}$, $q > 1$. Вместо одного параметра $r$ теперь
необходимо выбирать 3, однако, согласно численным экспериментам, сделать это проще, чем найти оптимальное значение $r$.
Интуитивно практическую эффективность предложенной схемы можно объяснить тем, что теперь работа метода происходит в
двух режимах: глобальный поиск при $r=r_{max}$ и локальный при $r=r_{min}$. Если во время фазы глобального посика
метод приблизился к глобальному минимуму, то во время следующей фазы оценка глобального минимума будет быстро уточнена.
Если двух фаз не хватило, то процесс продолжается. Таким образом выбирается более оптимальный trade-off between
exploration and exploitation. Далее будем обозначать метод, использующий описанную схему как AGS-AR.

\subsection{Other Optimization Methods}
\begin{itemize}
  \item \textbf{Multi Level Single Linkage} \cite{Kan1987StochasticGO}. MLSL is an improved
multistart algorithm.
  It samples low-discrepancy starting points and does local optimizations from them. In contrast to
the dummy multistart schemes
  MLSL uses some clustering heuristics to avoid multiple local descents to already explored local
minima.

@@

All the mentioned algorithms are available in source codes as parts of wide-spread optimization packages.
DIRECT, DIRECT$l$, CRS, MLSL and StoGO are part of the NLOpt library \cite{nlopt}.
Differential Evolution and DSA can be found in
the latest version of the SciPy \cite{scipy} package for Python.

@@

The results of various algorithms on different problem classes depend on the adjustments of
algorithms directly. In most cases, the authors of software implementations are oriented onto the
problems of medium difficulty. In order to obtain a satisfactory result when solving the essentially
multiextremal problems, a correction of some parameters is required. When conducting the
comparison, the following parameters for the methods were employed:
\begin{itemize}
  \item in the AGS-AR method, the parameter of alternation the
  global and local stages $q$ was set to be equal to $100\cdot N^2$, also $r_{min}=3,\:r_{max}=2\cdot r_{min}$;
  \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
  \item in the SDA method, the parameter \(visit=2.72\).
\end{itemize}

The rest parameters were varied subject to the problem class (see Table \ref{tab:params}).
For the AGS the value ot the $r$ parameter, such that the method solves all problems and performs the minimum amount of trials,
was estimated by brute force on the uniform grid with step $0.1$.

@@

The results of running the optimization methods on the considered problem classes are presented in
Tables \ref{tab:trials}, \ref{tab:solved}. The DIRECT, AGS and AGS-AR methods have demonstrated the
best convergence speed on all classes, at that AGS-AR inferior to DIRECT on the 2d problems from the
Simple classes and has an advantage on the problems of the Hard classes. As one can see from Table
\ref{tab:solved}, the deterministic methods (AGS, AGS-AR, DIRECT, and DIRECT\(l\)) were the
most reliable. Among the stochastic methods, MLSL and SDA have demonstrated the highest
reliability.

\begin{table}
\begin{center}
\caption{Averaged number of trials executed by optimization methods for solving the test
optimization problems}
\resizebox{\textwidth}{!}{%
  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
    \hline
    & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
  \hline
  \(F_{GR}\)     & 193.1 &  \textbf{158.3} & 400.3 & 182.3 & 214.9 & 947.2 & 691.2 & 1257.3 &
1336.8 \\
  \hline
  GKLS 2d Simple &  254.9 & 217.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2
& 1251.5 \\
  \hline
  GKLS 2d Hard   &  728.7 & \textbf{488.0} & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 &
1041.1 & 2532.2 \\
  \hline
  GKLS 3d Simple &  1372.1 & 1195.3 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 &
5956.94 & 3856.1 \\
  \hline
  GKLS 3d Hard   &  3636.1 & \textbf{1930.5} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 &
6914.3 & 7843.2 \\
  \hline
  GKLS 4d Simple &  26654.1 & 11095.7 & 37436.8 & \textbf{7824.3} & 15994.1 & 41514.3 &
21417.9 & 19157.7 & 59895.4 \\
  \hline
  GKLS 4d Hard   &  54536.8 &  \textbf{23167.8} &  73779.3 &  23204.4 &  54489.9 &  80247.2
&  68815.5 &  27466.1 &  109328.1  \\
  \hline
  GKLS 5d Simple &  29810.0 & 11529.0 & 143575.0 & \textbf{7166.5} & 13970.5 & 52647.6 &
34255.3 & 73074.5 & 91580.4 \\
  \hline
  GKLS 5d Hard   &  113129.1 & 67652.7 & 165192.8 & \textbf{66327.4} & 164390.6 & 138766.2
& 116973.1 & 105496.9 & 155123.8 \\
  \hline
\end{tabular}}
  \label{tab:trials}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Number of test optimization problems solved by methods}
  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
    \hline
    & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
  \hline
  \(F_{GR}\)     &  100 & 100 & 76  & 100 & 100 & 97  & 96  & 96  & 67\\
  \hline
  GKLS 2d Simple &  100 & 100 & 85  & 100 & 100 & 100 & 100 & 98  & 90\\
  \hline
  GKLS 2d Hard   &  100 & 100 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
  \hline
  GKLS 3d Simple &  100 & 97  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
  \hline
  GKLS 3d Hard   &  100  & 99   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
  \hline
  GKLS 4d Simple &  100 & 100 & 46  & 100 & 100 & 94  & 78  & 59  & 16 \\
  \hline
  GKLS 4d Hard   &  100 & 100 & 47  & 99  & 97  & 94  & 72  & 32  & 10  \\
  \hline
  GKLS 5d Simple &  100 & 100 & 68  & 100 & 100 & 98  & 100 & 77  & 9  \\
  \hline
  GKLS 5d Hard   &  97  & 99  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
  \hline
  \end{tabular}
  \label{tab:solved}
\end{center}
\end{table}

@@

In the present paper, several sequential and parallel global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
making the following conclusions:
\begin{itemize}
  \item the proposed modification of the stock AGS substantially outperforms it on the most complicated problems
  from the generated test set and allows to pay less attention to initial hyperparameter tuning.
  \item AGS-AR method  has demonstrated the convergence
speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems;
\end{itemize}
